% $Id: design_choices.tex,v 1.237 2003/10/24 10:24:24 mkomu Exp $

%%%%%%%%%%%%%%%%%%%% GENERAL TODO SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% - Put all of the figs into ``fig'' directory
% - Merge Torsten's comments in this paper
%   - Long sections should be split into smaller
%     - Example: \textbf{Keyword} begins this section..
%   - Some sections contain repeated information, is this bad?
%     - Resolver architecture contains redundant information
%       - Network socket API: The motivation behind the flags...
%       - If both transparency and AI_HIP...
%     - Userspace network applications
%        - HIP aware applications may require...
%        - Technically this is a correct way...
% - Jouni Korhonen suggested
%   - more stuff to DNS part -> Miika's thesis corrects this.
%   - more punctuality with the solutions (state/sequence transition graphs)
%   - separate ``we did not implement x because'' things from the text
%     because they are being repeated all over the text
%   - clarify the things we did not implement
%   - clarify why we implemented x with method y
% - mention about transport/tunnel mode and which/why our implementation
%   uses it
% - Merge Henderson's email comments in this paper
%   - One packet buffer - compare to Jukka's thoughts
%   - HIs can be distributed using CERTs - requires PKI
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,titlepage]{article}
\usepackage{fancyhdr}
\usepackage{a4wide}
%\usepackage{alltt}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{supertabular}
\usepackage{array}          % XX TODO: is this needed?
%\usepackage{enumerate}
\usepackage{varioref}
\usepackage{hyperref}       % keep this as the last one
\begin{document}

\input{design_choices_macros}

\title{Host Identity Protocol for Linux}
\author{Catharina Candolin, \texttt{candolin@tml.hut.fi} \\
        Miika Komu, \texttt{miika.komu@hiit.fi} \\
	Mika Kousa, \texttt{mika.kousa@hiit.fi} \\
	Janne Lundberg, \texttt{jlu@tcs.hut.fi} \\
	Kristian Slavov, \texttt{kristian.slavov@hiit.fi} \\
	\\
        Helsinki Institute for Information Technology \\       
        Helsinki University of Technology: \\
        Department of Computer Science and Engineering}
\maketitle


\fancyhf{} % Clear all fields
\fancyhead[C]{\small Host Identity Protocol for Linux}
\fancyfoot[C]{ \thepage }
\pagestyle{fancy}

\tableofcontents

\newpage

\section*{Acknowledgements}

This research is funded by the National Technology Agency of Finland
(TEKES), Elisa Communications, Ericsson, Nokia, TeliaSonera, Creanor,
and More Magic Software. We thank Jukka Ylitalo, Jorma Wall, Petri
Jokela, and especially Pekka Nikander from Ericsson Research for
fruitful cooperation and interoperability testing with their
implementation of HIP for BSD. Furthermore, we are grateful for the
valuable discussions and comments we have had with Andrew McGregor,
Thomas Henderson, and Jeff Ahrenholz. Torsten R{\"u}ger (HIIT) and
Jouni Korhonen (TeliaSonera) gave also some valuable comments on this
document.

\newpage

%%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
One of the main problems with IP has been its lack of
security. Although IPSec and DNSSec have provided some level of
security to IP, the notion of a true identity for hosts is still
missing. Typically, the IP address of the host has been used as the
host identity, regardless of the fact that it is nothing more than
routing information. The purpose of the Host Identity Payload/Protocol
(HIP) architecture is to add a cryptographically based name space, the
Host Identity, to the IP protocol. The Host Identity serves as the
identity of the host, whereas the IP address is merely used for
routing purposes.  In this document, we describe the HIP architecture
further, and present our IPv6 based implementation of HIP for Linux.
\end{abstract}

\input{hipl.tex}

\bibliographystyle{plain}
\bibliography{design_choices}


\appendix
\newpage
\pagebreak

\include{catalogue}

\end{document}
